Working title, The impact of heterospecific pollen on plant reproductive success is mediated by phylogenetic distance and floral reproductive traits

### 2018-11-30 Manuscript_draft/Manuscript_draft.Rmd
reading and writing day

Possible journals to publish: New phytologist, journal of ecology, oikos...

@@

In natural systems plant species normally coexist and share their floral visitors with other species @bascompte2003. This pollinator sharing from the plant perspective at the pre-pollination stage can be negative due to competition @pauw2013 or positive due to facilitation @carvalheiro2014. Once, the floral visitor has arrived to the flower, pollen deposition on the stigma can take place and hence ovule fertilization. An increasing number of visits generally correlates with higher chances of fertilization (Refs). However this is not always the case, among these possible flower visitors we find also nectar robbers and pollen thiefs (Refs) and the quality of pollen that is deposited on the stigma is also highly relevant to the pollination succes (Refs). Moreover, other less study issues in the pollination process are conspecific pollen loss and the arrival of foreign pollen which can have important detrimental effects for the species fitness @morales2008 (More refs).

@@

Recent important studies have advanced in the ecological importance of heterospecific pollen effect @morales2008 @ashamn2013 @arceo2016. A general overview of it can reduce species fitness (REFS) but not in al the cases (Refs). Other general facts... Quantity of foreign pollen that arrives...

>>>>

Pollinator sharing can have negative consequences for species fitness with the arrival of foreign pollen. However, the costs of heterospecific pollen are not yet well understood. For this reason, we have conducted a glasshouse experiment where we try to understand how phylogenetic relatedness and the different traits of these species are involved in this process. We experimentally crossed 10 species belonging to three different families: Brassicaceae, Solanaceae and Convolvulaceae. Overall, more than 4000 crosses were done and seed set and pollen tubes were considered as proxy of effect. We found that for all species foreign pollen (50% or less) reduced seed set. Moreover, the seed set reduction is not dependent on the degree of relatedness of the pollen donor. However, the effect is governed by the degree of relatedness and the traits of the species recipient. Our results show that the outcome of heterospecific pollen deposition is determined in greater degree by the traits of the pollen recipient than the pollen donor and that certain traits such as compatibility system are crucial to understand the costs of heterospecific pollen.

@@

In natural systems plant species normally coexist and share their floral visitors with other species @bascompte2003. This pollinator sharing from the plant perspective at the pre-pollination stage can be negative due to competition @pauw2013 or positive due to facilitation @carvalheiro2014. Once the floral visitor has arrived to the flower, pollen deposition on the stigma can take place and hence ovule fertilization. An increasing number of visits generally correlates with higher chances of fertilization @engel2003. However this is not always the case, among these possible flower visitors we find also nectar robbers and pollen thiefs @inouye1980 and the quality of pollen that is deposited on the stigma is also highly relevant to the pollination succes @aizen2007. Moreover, other less study issues in the pollination process are conspecific pollen loss and the arrival of foreign pollen which can have important detrimental effects for the species fitness @morales2008 @ashman2013.

@@

Recent studies have advanced in the ecological importance of heterospecific pollen effect @morales2008 @ashamn2013 @arceo2016. A general overview of heterospecific pollen effect is that it can play an important role on species fitness but seems to be context dependent and not always fitness decreased is produced @morales2008. There are several traits that are postulated to be highly relevant to this process but the correlation with heterospecific pollen effect remains porrly understood.

Mention relatedness somwhere here.

Moreover, the different methods used make difficult to compare among different experiments. For instance many articles do not describe the foreign pollen quantity used to test for effect or how this pollen was applied (Refs). The small size of pollen, great environmental variability and population outcrossing rates among others make of this field of study a daunting task and higly unpredictable.

can reduce species fitness (REFS) but seems to be highly contex-dependent. There are hypothesized that some traits can play a crucial role in this species interaction such as stigma type, pollen size,

### 2018-11-30 Manuscript_draft/Manuscript_draft.tex
reading and writing day

Possible journals to publish: New phytologist, journal of ecology,
oikos\ldots{}

@@

In natural systems plant species normally coexist and share their floral
visitors with other species Bascompte et al. (2003). This pollinator
sharing from the plant perspective at the pre-pollination stage can be
negative due to competition Pauw (2013) or positive due to facilitation
Carvalheiro et al. (2014). Once, the floral visitor has arrived to the
flower, pollen deposition on the stigma can take place and hence ovule
fertilization. An increasing number of visits generally correlates with
higher chances of fertilization (Refs). However this is not always the
case, among these possible flower visitors we find also nectar robbers
and pollen thiefs (Refs) and the quality of pollen that is deposited on
the stigma is also highly relevant to the pollination succes (Refs).
Moreover, other less study issues in the pollination process are
conspecific pollen loss and the arrival of foreign pollen which can have
important detrimental effects for the species fitness Morales and
Traveset (2008) (More refs).

@@

Recent important studies have advanced in the ecological importance of
heterospecific pollen effect Morales and Traveset (2008)
({\textbf{???}}) Arceo-Gómez and Ashman (2016). A general overview of it
can reduce species fitness (REFS) but not in al the cases (Refs). Other
general facts\ldots{} Quantity of foreign pollen that arrives\ldots{}

@@

\hypertarget{refs}{}
\hypertarget{ref-Allaire_2018}{}
Allaire, J., Y. Xie, J. McPherson, J. Luraschi, K. Ushey, A. Atkins, H.
Wickham, J. Cheng, and W. Chang. 2018. Rmarkdown: Dynamic documents for
r.

@@

@@

>>>>

Pollinator sharing can have negative consequences for species fitness
with the arrival of foreign pollen. However, the costs of heterospecific
pollen are not yet well understood. For this reason, we have conducted a
glasshouse experiment where we try to understand how phylogenetic
relatedness and the different traits of these species are involved in
this process. We experimentally crossed 10 species belonging to three
different families: Brassicaceae, Solanaceae and Convolvulaceae.
Overall, more than 4000 crosses were done and seed set and pollen tubes
were considered as proxy of effect. We found that for all species
foreign pollen (50\% or less) reduced seed set. Moreover, the seed set
reduction is not dependent on the degree of relatedness of the pollen
donor. However, the effect is governed by the degree of relatedness and
the traits of the species recipient. Our results show that the outcome
of heterospecific pollen deposition is determined in greater degree by
the traits of the pollen recipient than the pollen donor and that
certain traits such as compatibility system are crucial to understand
the costs of heterospecific pollen.

@@

In natural systems plant species normally coexist and share their floral
visitors with other species Bascompte et al. (2003). This pollinator
sharing from the plant perspective at the pre-pollination stage can be
negative due to competition Pauw (2013) or positive due to facilitation
Carvalheiro et al. (2014). Once the floral visitor has arrived to the
flower, pollen deposition on the stigma can take place and hence ovule
fertilization. An increasing number of visits generally correlates with
higher chances of fertilization Engel and Irwin (2003). However this is
not always the case, among these possible flower visitors we find also
nectar robbers and pollen thiefs Inouye (1980) and the quality of pollen
that is deposited on the stigma is also highly relevant to the
pollination succes Aizen and Harder (2007). Moreover, other less study
issues in the pollination process are conspecific pollen loss and the
arrival of foreign pollen which can have important detrimental effects
for the species fitness Morales and Traveset (2008) Ashman and
Arceo-Gómez (2013).

@@

Recent studies have advanced in the ecological importance of
heterospecific pollen effect Morales and Traveset (2008)
({\textbf{???}}) Arceo-Gómez and Ashman (2016). A general overview of
heterospecific pollen effect is that it can play an important role on
species fitness but seems to be context dependent and not always fitness
decreased is produced Morales and Traveset (2008). There are several
traits that are postulated to be highly relevant to this process but the
correlation with heterospecific pollen effect remains to be tested.
Moreover, the different methods used make difficult to compare among
different experiments. For instance many articles do not describe the
foreign pollen quantity used to test for effect or how this pollen was
applied (Refs). The small size of pollen, great environmental
variability and population outcrossing rates among others make of this
field of study a daunting task and higly unpredictable.

can reduce species fitness (REFS) but seems to be highly
contex-dependent. There are hypothesized that some traits can play a
crucial role in this species interaction such as stigma type, pollen
size,

@@

\hypertarget{refs}{}
\hypertarget{ref-aizen2007}{}
Aizen, M. A., and L. D. Harder. 2007. Expanding the limits of the
pollen-limitation concept: Effects of pollen quantity and quality.
Ecology 88:271--281.

\hypertarget{ref-Allaire_2018}{}
Allaire, J., Y. Xie, J. McPherson, J. Luraschi, K. Ushey, A. Atkins, H.
Wickham, J. Cheng, and W. Chang. 2018. Rmarkdown: Dynamic documents for
r.

@@

\hypertarget{ref-ashman2013}{}
Ashman, T.-L., and G. Arceo-Gómez. 2013. Toward a predictive
understanding of the fitness costs of heterospecific pollen receipt and
its importance in co-flowering communities. American Journal of Botany
100:1061--1070.

@@

\hypertarget{ref-engel2003}{}
Engel, E. C., and R. E. Irwin. 2003. Linking pollinator visitation rate
and pollen receipt. American Journal of Botany 90:1612--1618.

\hypertarget{ref-inouye1980}{}
Inouye, D. W. 1980. The terminology of floral larceny. Ecology
61:1251--1253.
